---
title: "Find Adult Cornea Markers"
author: David McGaughey
date: 2020-02-10
output:  
  html_notebook:
    theme: flatly
    toc: true
    toc_float: true
---

# Load libraries, import count data and metadata

The `gene_tpm` import will take some time, as this file is 614mb
```{r}
library(tidyverse)
#https://eyeintegration.nei.nih.gov -> Data -> Data Download for links
metadata <- read_tsv('https://hpc.nih.gov/~mcgaugheyd/eyeIntegration/2019_metadata_04.tsv.gz')
gene_tpm <- read_tsv('https://hpc.nih.gov/~mcgaugheyd/eyeIntegration/2019_gene_TPM_04.tsv.gz')
```

# Let's see what we have
```{r}
metadata %>% sample_n(10)

gene_tpm[1:10,c(1,100:110)]
```

# We can aggregate data to the sub_tissue level in a few steps.

## First make the data *long* with `pivot_longer`
This makes it easier for the computer to work with the data
```{r}
gene_tpm_long <- gene_tpm %>% 
  pivot_longer(cols=2:ncol(gene_tpm), names_to = 'sample_accession', values_to = 'TPM')

gene_tpm_long %>% sample_n(10)
```

## Now `left_join` to label each sample with the tissue and sub_tissue fields
```{r}
gene_tpm_long <- left_join(gene_tpm_long, 
                           metadata %>% select(sample_accession, Tissue, Sub_Tissue), 
                           by = 'sample_accession')

gene_tpm_long %>% sample_n(10)
```

## Now we can aggregate expression to the `Sub_tissue` level 
We will take the mean gene (`ID`) expression of each individual sample, by the `Sub_Tissue` assignment
```{r}
gene_tpm_long_by_sub_tissue <- gene_tpm_long %>% 
  group_by(ID, Sub_Tissue) %>% summarise(TPM = mean(TPM)) 

gene_tpm_long_by_sub_tissue %>% 
  sample_n(10)
```

# Learning stuff now!
## 3 most expressed genes in the Cornea Sub Tissues
```{r}
gene_tpm_long_by_sub_tissue %>% 
  filter(grepl('Cornea', Sub_Tissue)) %>% 
  group_by(Sub_Tissue) %>% 
  top_n(3, wt = TPM) %>% 
  mutate(TPM = log2(TPM + 1)) %>% 
  arrange(Sub_Tissue)
```

## Genes double in expression with Cornea - Adult Tissue against most other `Sub_Tissue`

### First, how many `Sub_Tissues` do we have?

`r gene_tpm_long_by_sub_tissue$Sub_Tissue %>% unique() %>% length()`
```{r}
gene_tpm_long_by_sub_tissue$Sub_Tissue %>% unique() %>% length()
```
`Count` is the number of (sub) tissues that the gene expression is 1 log2(TPM + 1) (double, or two fold) higher in `Cornea - Adult`. The max we can have is 74. 

`Mean Delta log2(TPM+1)` is the average (mean) delta fold change between Cornea - Adult and the other tissues. Each log2(TPM+1) is a doubling in expressed.

The `Mean Cornea - Adult log2(TPM + 1)` is the average `Cornea - Adult` expression.
This column is useful to remove genes with overall low expression. 

We see that we have over 400 genes that are:
  
  1. More than double in expression than most (70 or more) of the other `Sub_Tissues`
  2. Expressed in `Cornea - Adult` above log(TPM + 1) > 10 (which we use as a rough "highly expressed" cut-off)

```{r}
left_join(gene_tpm_long_by_sub_tissue %>% 
            filter(Sub_Tissue == 'Cornea - Adult Tissue') %>% rename(cTPM = TPM),
          gene_tpm_long_by_sub_tissue %>% 
            filter(!Sub_Tissue == 'Cornea - Adult Tissue'),
          by = c('ID')) %>% 
  mutate(Delta = log2(cTPM + 1) - log2(TPM + 1)) %>% 
  filter(Delta > 1) %>% 
  group_by(ID) %>% 
  summarise(Count = n(), `Mean Delta log2(TPM+1)` = mean(Delta), `Mean Cornea - Adult log2(TPM + 1)` = mean(log2(cTPM + 1))) %>% 
  arrange(-Count, -`Mean Delta log2(TPM+1)`) %>% 
  filter(Count >= 70,
         `Mean Cornea - Adult log2(TPM + 1)` > 10) %>% 
  DT::datatable()
```

# Exercise(s) for the reader
What about genes in other Cornea sub tissues?

Genes that are LOWER in expression in cornea?

# Session Info
```{r}
devtools::session_info()
```
