---
title: "Fetal and Organoid Human Retina"
author: David McGaughey
date: 2019-01-16
output:  
  html_notebook:
    theme: flatly
    toc: true
    toc_float: true
---

# Load libraries, connect to 2019 SQLite DB
```{r}
library(pool)
library(tidyverse)
library(RSQLite)
library(ggrepel)

# edit this to match the path of the 2019 Eye on a Disk dataset (EOAD - "E Ode"?)
# you can get this database here:
# https://github.com/davemcg/eyeIntegration_app
pool <- dbPool(drv = SQLite(), dbname = "/Volumes/data/projects/nei/mcgaughey/auto_eyeIntegration/results/eyeIntegration_human_expression_2019_v100.sqlite")
```

# Tables in 2019 dataset
```{r}
pool %>% dbListTables() %>% as_tibble()
```
# Load in t-SNE coordinates (and attached metadata)
For perplexity 50
```{r}
tsne_50 <- pool %>% tbl('tSNE_bulk_RNA') %>% 
  filter(perplexity == 50) %>% as_tibble()
tsne_50 %>% colnames() %>% enframe()
```

# Plot Eldred Organoid Retina against all other samples
```{r}
tsne_50 %>% mutate(Organoid = case_when(grepl("Organoid", Sub_Tissue) ~ "Organoid",
                                        TRUE ~ "Not Organoid")) %>% 
  ggplot(aes(x=X1, y=X2, colour = Organoid)) +
  geom_point() + theme_minimal()
```

# Now just Retina
```{r}
tsne_50 %>% 
  mutate(Organoid = case_when(grepl("Organoid", Sub_Tissue) ~ "Organoid",
                              TRUE ~ "Not Organoid")) %>% 
  filter(Tissue == 'Retina') %>% 
  ggplot(aes(x=X1, y=X2, colour = Organoid, shape = Sub_Tissue)) +
  geom_point() + theme_minimal()
```

# Now show the ages of the Organoids
We see the little outlier group are age 10 and 20 day organoids. Everything older
looks very "retina" like (clusters closely to the fetal and stem cell based retina).
```{r}
tsne_50 %>% 
  mutate(Organoid = case_when(grepl("Organoid", Sub_Tissue) ~ "Organoid",
                              TRUE ~ "Not Organoid")) %>% 
  filter(Tissue == 'Retina') %>% 
  mutate(sample_attribute = case_when(grepl("Organoid", Sub_Tissue) ~ sample_attribute,
                                      TRUE ~ "")) %>% 
  rowwise() %>% 
  mutate(Age = as.numeric(str_split(sample_attribute, '_')[[1]][2])) %>% 
  ggplot(aes(x=X1, y=X2, colour = as.factor(Age), shape = Sub_Tissue)) +
  geom_jitter() + theme_minimal() 
```

# What are the Day 10 and 20 organoid grouped with?
They are near (undifferentiated) human stem cells and some
lens stem cells (which I would guess are not well differentiated?).

So it appears that in the Eldred/Johnston differentiation process, the
"retina-ness" shift is happening between 20-35 days.

Another shift happens between 69 and 111 days (look at above plot).
```{r}
tsne_50 %>% 
  mutate(Organoid = case_when(grepl("Organoid", Sub_Tissue) ~ "Organoid",
                              TRUE ~ "Not Organoid")) %>% 
  filter(X1 > 29 & X1 < 33) %>% 
  mutate(sample_attribute = case_when(grepl("Organoid", Sub_Tissue) ~ sample_attribute,
                                      TRUE ~ "")) %>% 
  rowwise() %>% 
  mutate(Age = as.numeric(str_split(sample_attribute, '_')[[1]][2])) %>% 
  ggplot(aes(x=X1, y=X2, colour = as.factor(Age), shape = Sub_Tissue)) +
  geom_point() + theme_minimal() 
```

