---
title: "eyeIntegration Viewing Stats"
output: html_notebook
---

Pull in IP addresses, filter to only keep IP addresses over 11 views (I think any 11 or below is a bot, since they are so common)

`eyeIntegration.stats` is generated as follows:

1. Select all `Daily_eyeinegration_frequent_IP_access_list` emails in Outlook
2. Drag all emails to empty folder
3. Run `cat *eml | grep "^[[:digit:]]" > eyeIntegration.stats`

```{r}

library(tidyverse)
library(rworldmap)
library(ipstack)
library(lubridate)
library(patchwork)

eml_files <- list.files('~/data/eyeIntegration_eml_emails_up_to_2021_12_24/test', full.names = TRUE, pattern = '*eml', recursive = TRUE)
eml_reader <- function(eml){
  raw <- read_lines(eml)
  
  tsv <- read_delim(eml, col_names = c('ip','Page Views'), delim = ' ')
  tsv <- tsv %>% filter(grepl('^\\d', ip)) %>% mutate(`Page Views` = trimws(`Page Views`) %>% as.integer())
  tsv$`Page Views` <- gsub('<br>', '', tsv$`Page Views`)
  
  chunk <- grep('Date', raw,value = TRUE) %>% substr(., 12,22) %>% str_split(., ' ', simplify = TRUE)
  date <- lubridate::mdy(paste(chunk[2], chunk[1], chunk[3]))
  
  tsv$date <- date
  
  tsv 
  
}

eml_list <- list()
for (i in eml_files){
  eml_list[[i]] <- eml_reader(i)
}

ipaddresses <- eml_list %>% bind_rows() %>% 
  mutate(`Page Views` = as.numeric(`Page Views`)) %>%  
  filter(`Page Views` > 11)

# if loading previous work
load('visitor_stats_2021_12.Rdata')
```

Total page view from May 2017 to Present
```{r}
sum(ipaddresses$`Page Views`)
```

Chart of page views increasing over time
```{r, fig.width=2.5}
a <- ipaddresses %>% 
  filter(`Page Views` > 11, date < as.Date('2021-12-01')) %>% 
  group_by(date) %>% 
  summarise(`Visitor Count` = length(unique(ip))) %>% 
  mutate(`Visitor Count` = cumsum(`Visitor Count`)) %>% 
  ggplot(aes(x=date,y=`Visitor Count`)) + 
  geom_line() +
  cowplot::theme_cowplot() +
  xlab('Date') + ylab('eyeIntegration\nCumulative\nVisitor Count') +
  expand_limits(x = as.Date(c("2017-05-01", "2022-01-01"))) +
  theme(text = element_text(size = 25))    

a
```   

Get latitude and longitude for all unique IPs (takes a few minutes to query)
```{r}
unique_IP <- ipaddresses$ip %>% unique()
#ip_info <- list()

# ipstack (useful, but no ISP info)
# for (i in unique_IP){
#   ip_info[[i]] <- ip_lookup(i, ipstack_api_key = '7d06f26c4f0f61bd9edd20df0219f874', hostname = T)
# }

for (i in unique_IP){
  if (!(i %in% names(ip_info))){
    print(i)
    ip_info[[i]] <- ipapi::geolocate(i, .progress = FALSE)
    Sys.sleep(2)
  }
}

ip_tibble <- ip_info %>% bind_rows()
```

Save
```{r}
save(ip_tibble, ip_tibble, ipaddresses, ip_info, file = '~/git/eyeIntegration_app/visitor_stats_2021_12.Rdata')
```

Interactive Map
```{r}
# library(ggiraph)
# world <- map_data("world")
# world <- world[world$region != "Antarctica",]

#world <- c(geom_polygon(aes(long,lat,group=group), size = 0.1, colour= "#090D2A", fill="white", alpha=0.8, data=worldmap))
# code <- ggplot() + geom_map(data=world, map=world,
#                     aes(x=long, y=lat, map_id=region),
#                     color="#090D2A", fill="white", size=0.05, alpha=1/4) + 
#   geom_point_interactive(aes(x = geo_locate$longitudes, y = geo_locate$latitudes, tooltip=geo_locate$cities)) + theme_void() 
# ggiraph(code = print(code) )
```

Regular Map
```{r, fig.width = 6, fig.height=2}
world <- map_data("world")
world <- world[world$region != "Antarctica",]
ggplot(data = ip_tibble, aes(x = lon, y = lat, colour = country)) + 
  geom_map(data=world, map=world,
           aes(x=long, y=lat, map_id=region),
           color="#090D2A", fill="lightgray", size=0.05, alpha=1/4) + 
  geom_point() + 
  theme_void() +
  scale_color_viridis_d()

```

# Figure
```{r, fig.width = 7, fig.height=6}
isp_location <- ipaddresses %>% left_join(ip_tibble, by = c('ip' = 'query')) %>% group_by(isp) %>% summarise(Usage = sum(`Page Views`), lat = mean(lat), lon = mean(lon))

isp_location_academic <- isp_location %>%
  filter(grepl('Agency|Acad|Scie|Centr|Center|Schoo|Univ|Hosp|Stat|Colle|Resea|Publ|Nation|Instit|Medici', isp), 
         !grepl('BT Public', isp)) %>% 
  arrange(-Usage)

c <- isp_location_academic %>% 
  head(25) %>%
  mutate(isp = factor(isp, levels = rev(isp_location_academic$isp))) %>%  
  ggplot(aes(x=isp, y = Usage)) + 
  geom_bar(stat = 'identity', fill = 'darkblue') +  
  ggbreak::scale_y_break(c(35000, 240000)) +
  scale_y_continuous(breaks = c(5000, 15000, 25000, 240000)) +
  coord_flip() +
  cowplot::theme_cowplot() +
  xlab('Top 25\nResearch Users') +
  ylab('Page Views') +
  theme(text = element_text(size = 20)) 



b <- ggplot(data = isp_location, aes(x = lon, y = lat) ) + 
  geom_map(data=world, map=world,
           aes(x=long, y=lat, map_id=region),
           color="#090D2A", fill="lightgray", size=0.05, alpha=1/4) + 
  geom_point(color = 'black', size = 4, alpha = 0.5) + 
  geom_point(data = isp_location_academic, color = 'darkblue', size = 5, alpha = 0.7) +
  theme_void() +
  scale_color_viridis_c()

first <- (a + c)  + 
  plot_layout(widths = c(1, 2)) +
  plot_annotation(theme = theme(text = element_text(size = 25)))

first / b +
  plot_layout(heights = c(1, 1.4))  +
  plot_annotation(
    title = 'eyeIntegration Worldwide Usage',
    subtitle = glue::glue('Users from {ip_tibble$country %>% unique() %>% length()} countries, {ip_tibble$city %>% unique() %>% length()} cities, and {isp_location_academic %>% nrow} research institutions.'),
    caption = 'Blue points are research instituations',
    theme = theme(text = element_text(size = 25))
  )
```
